---
title: "Automating web scraping with R"
author: Alex Sanchez and Francesc Carmona
institute: Genetics Microbiology and Statistics Department <br> Universitat de Barcelona
date: October 2022 # "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, "css4WScourse.css" ]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: true
editor_options: 
  chunk_output_type: console
---

<style type="text/css">
.remark-slide-content {
    font-size: 28px;
    padding: 1em 2em 1em 2em;
}
.left-code {
  color: #777;
  width: 38%;
  height: 92%;
  float: left;
}
.right-plot {
  width: 60%;
  float: right;
  padding-left: 1%;
}
<style>
span.small {
  font-size: smaller;
}
</style>

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, echo=TRUE,
        message=FALSE,warning=FALSE,
        fig.dim=c(4.8, 4.5), fig.retina=2, out.width="100%")

knitr::opts_chunk$set(echo = TRUE)

knitr::knit_hooks$set(mysize = function(before, options, envir) {
  if (before) 
    return(options$size)
})
```

# Outline

.columnwide[
  ### 1) [Introduction](#Introduction)
  ### 2) [User defined functions](#functions)
  ### 3) [Changing the execution flow](#loops)
  ### 4) [References and Resources](#Resources)
]

---

class: inverse, middle, center

name: Introduction

# Introduction

---

# Introduction

- We have introduced R  as a *a language (a tool), to manage and analyze data*.
- It is also a *programming language*
  + It is simple and versatile
  + The user can create new functions that adapt to their
needs
  + It is widely used (2nd most widely used in Data Science)
  + Users provide the community with a high variety of solutions ("packages")
  + As a programming language it is not, however, very efficient
    
---

# Example 1: Why we need programming

- It is very common that one has to do repetitive tasks on a the same type of datasets, e.g. 
  - Data produced periodically o,
  - Data from multiple sources but have the same structure.

- For example, given a file with information about the cities of a given province, we are required to produce a simpler version:
  - With less columns
  - Without spaces or accents in the column names
  - With the appropriate data types for each column

---

# Example 1: Transformations 

<div style="font-size:16pt">

```{r}
library(dplyr); library(janitor)
MunicipisBARC <- read.csv("datasets/MunicipisBARC.csv", quote='"')
MunicipisBreu <- MunicipisBARC %>% select (2,9,10,27,28) %>%  
  janitor::clean_names() %>% 
  mutate (across(c("codi_de_comarca", "nom_de_la_comarca"), as.factor))
```

---

# Example 1: Before and after transform

 <div style="font-size:12pt">

.pull-left[
```{r}  
summary(MunicipisBARC[,c(2,9,10,27,28)])
```
]

.pull-right[

```{r}
summary(MunicipisBreu)
```

`codi_de_comarca` and `nom_de_la_comarca` are now factors.
]

---

# Repeating the transformation

- How should we proceed if these changes had to be applied repeteadly to many distinct files (with same structure)

- One solution may consists of:
  - providing some way to encapsulate all steps needed to do the transformation 
  - in such a way that they can be easily applied to a file everytime they are required.
  
- This is an example of a "function" (a type of sub-program in R) that we would use to automate this preprocessing.

---

class: inverse, middle, center

name: functions

# User defined functions

---

# Functions are named expressions

- A function is a set of statements organized together to perform a specific task. 
- R has a large number of in-built functions.
- Users can create their own functions, for those situations where they wish to apply the same set of instructions more than once.

<div style="font-size:18pt">
```{r eval=FALSE}
function_name <- function(arg_1, arg_2, ...) {
   sentence 1
   ...
   sentence n
   return(result)
}
```

Go [here](https://www.tutorialspoint.com/r/r_functions.htm) for more information on functions.

---
# A preprocessing function

We can encapsulate preprocessing in a function:
<div style="font-size:14pt">
```{r}
preprocessa <- function(nomArxiuMunicipi){
  require(dplyr); require(janitor)
  unMunicipi <- read.csv(nomArxiuMunicipi)
  unMunicipiBreu <- unMunicipi %>% select (2,9,10,27,28) %>%  
  janitor::clean_names() %>% 
  mutate (across(c("codi_de_comarca", "nom_de_la_comarca"), as.factor))
  return(unMunicipiBreu)
}
```

And use it whenever is required. Assuming we had the files "MunicipisBARC.csv", "MunicipisGIRO.csv" and "MunicipisLLEI.csv" available we would do: 

```{r eval=FALSE}
BCNBreu <- preprocessa("MunicipisBARC.csv")
GiroBreu <- preprocessa("MunicipisGIRO.csv")
LleidaBreu <- preprocessa("MunicipisLLEI.csv")
```

---

# Scraping a recipes site

- Imagine we are scraping a recipes site
- The code below extracts (without cleaning it) a recipe for brownies.

<div style="font-size:12pt">
```{r}
library(rvest)
brownies <- read_html("https://www.allrecipes.com/recipe/25080/mmmmm-brownies/")
ingredients <- brownies %>% 
  html_elements( "#mntl-structured-ingredients_1-0") %>% 
  html_text2() %>% stringr::str_split("\\n\\n")

xpath4Directions <- '//*[(@id = "recipe__steps_1-0")]'

directions <-  brownies %>% 
  html_elements( xpath=xpath4Directions) %>% 
  html_text2() %>% stringr::str_split("\\n\\n") 

```

</div>
- Selectors were obtained inspecting the page source code with SelectorGadget or Google Developper tools.

---
# The scraped recipe

<div style="font-size:12pt">
.pull-left[
```{r}
show(ingredients)
```
]
.pull-right[
```{r}
show(directions)
```
]
---

# A function to scrape recipes

- Proceed similarly as before:
  - abstracting the process and 
  - turning what is different every time (URL) into arguments
  
<div style="font-size:12pt">
```{r}
scrape_recipes <- function(URL) {
  aDessert <- read_html(URL)
  ingredients <- aDessert %>% 
    html_elements( "#mntl-structured-ingredients_1-0") %>% 
    html_text2() %>% stringr::str_split("\\n\\n")

  xpath4Directions <- '//*[(@id = "recipe__steps_1-0")]'

  directions <-  aDessert %>% 
    html_elements( xpath=xpath4Directions) %>% 
    html_text2()%>% stringr::str_split("\\n\\n")
  return(list(Ingredientes=ingredients, Receta=directions))
}
```


---
# The scraped recipe (2)

<small>
```{r}
library(rvest)
recipeURL <- "https://www.allrecipes.com/recipe/25080/mmmmm-brownies/"
brownies <- scrape_recipes (recipeURL)
```
</small>
<div style="font-size:12pt">
.pull-left[
```{r}
show(brownies[["Ingredientes"]])
```
]
.pull-right[
```{r}
show(show(brownies[["Receta"]]))
```
]

---

class: inverse, middle, center

name: flowchange

# Changing the flow

---

# Changing the flow of execution


- R, as most ordinary programming languages, is executed lineally, that is from the first to last line.
- Sometimes this needs to be changed.
    + Taking alternative flows according to certain conditions
    + Repeating some instructions while certain condition holds, or a fixed number of times,...
    
- This can be acomplished using *Flow Control Structures*

---

# Loop controlled by a counter: `for`

- Loops are used in programming to repeat a specific block of code made by one or more instructions.
- Syntax of `for` loops:

```{r eval=FALSE}
    for (val in sequence)
    {
    statement
    }
```

- `sequence` is a vector and `val` takes on _each of its values_ during the loop. 
- In each iteration, `statement` is evaluated.

---

# Example of `for` loop

- A `for` loop can be used to preprocess a list of selected files
- Assume we have the list of four files to be processed, and __*we know they have the same structure*__.
- To process them all in one step do (not run):

<div style="font-size:15pt">
```{r eval=FALSE}
llistaMunicipis <- c("MunicipisBAR.csv", "MunicipisGIR.csv", 
                     "MunicipisLLE.csv", "MunicipisTAR.csv" )
for (nomFitxerMunicipis in llistaMunicipis) {
  municipisProvincia <- preprocessa("nomFitxerMunicipis")
  summary(municipisProvincia)
}
```
An alternative way to run the loop:
```{r eval =FALSE}
for (i in 1:length(llistaMunicipis) {
  municipisProvincia <- preprocessa(llistaMunicipis[i])

```

---

# Exercise

- Create a `for` loop that reads all filenames in your datasets directory (or the directory you decide) and prints the name of the file and the column names in the screen.

---

# Scraping multiple recipes

- Imagine we want to process not one but many desserts' recipes from the web 
"https://www.allrecipes.com/".
- This can be done using a simple for loop:


<div style="font-size:11pt">
```{r eval=FALSE}
recipe_urls <- c("https://www.allrecipes.com/recipe/25080/mmmmm-brownies/",
                 "https://www.allrecipes.com/recipe/27188/crepes/",
                 "https://www.allrecipes.com/recipe/22180/waffles-i/")
listOfRecipes <- list()
for (i in 1:length(recipe_urls)) {
  listOfRecipes[i] <- scrape_recipes(recipe_urls[i])
}
```
<div>

<div style="font-size:18pt">
- Notice that the resulting scraped recipes are now stored in a _list_ that will be eventually processed by the user.

---
# Exercise

- Write a simple function to print one recipes obtained using the function `scrape_recipes`

- Use this function to print all the recipes collected in the list "listOfRecipes" 
-