---
title: "Web Scraping Course - Examples & Exercises"
format:
  html:
    toc: true
    toc-depth: 3
editor: visual
---

# Ch 1: **Introduction to HTML and Web Scraping**

```{r}
library(rvest)
library(magrittr)
```

1.  Take the `html_excerpt_raw` variable and turn it into an HTML document that R understands using a function from the `rvest` package.

```{r}
html_excerpt_raw <- '
<html> 
  <body> 
    <h1>Web scraping is cool</h1>
    <p>It involves writing code – be it R or Python.</p>
    <p><a href="https://datacamp.com">DataCamp</a> 
		has courses on it.</p>
  </body> 
</html>'
# Turn the raw excerpt into an HTML document R understands
html_excerpt <- read_html(html_excerpt_raw)
html_excerpt
```

2.  Extract the `ol` node from the `list_html` document, using the singular version of the `rvest` function that can be used to query nodes.

    ```{r}
    list_raw_html <- '<ul>
    <li>Learn HTML</li>
    <li>Learn CSS</li>
    <li>Learn R</li>
    <li>Scrape everything!*</li>
    </ul>'
    # Read in the corresponding HTML string
    list_html <- read_html(list_raw_html)
    # Extract the ol node
    ol_node <- list_html %>% 
    	html_element('ol')
    ```

3.  Extract all the `a` nodes *that are within the bulleted list*, using `html_elements()`.

    ```{r}
    hyperlink_raw_html <- '<h3>Helpful links</h3>
    <ul>
      <li><a href="https://wikipedia.org">Wikipedia</a></li>
      <li><a href="https://dictionary.com">Dictionary</a></li>
      <li><a href="https://duckduckgo.com">Search Engine</a></li>
    </ul>'
    # Extract all the a nodes from the bulleted list
    links <- hyperlink_raw_html %>% 
      read_html() %>% 
      html_elements("a")
    ```

4.  Recover the table from the page: \[https://en.wikipedia.org/wiki/List_of_World_Heritage_in_Danger\](https://en.wikipedia.org/wiki/List_of_World_Heritage_in_Danger)

    ```{r eval=FALSE}
    aURL <- "https://en.wikipedia.org/wiki/List_of_World_Heritage_in_Danger"
    mainTable <- read_html(url) # %>% 
      rvest::html_table() %>% 
      rvest::html_text2()
      
        
    ```

5.  Assume you have stored the mountains table in an html object

<!-- -->

    | Mountain      | Height | First ascent | Country      |
    |:--------------|:-------|:-------------|:-------------|
    | Mount Everest | 8848   | 1953         | Nepal, China |
    | \...          |        |              |              |

Turn the `table` into a data frame called `mountains`.

    ```{r}
    mountains_html <-read_html("<table id='clean'>
    <tr>
      <th>Mountain</th>
      <th>Height</th>
      <th>First ascent</th>
      <th>Country</th>
    </tr>
    <tr>
      <td>Mount Everest</td>
      <td>8848</td>
      <td>1953</td>
      <td>Nepal, China</td>
    </tr>
    <tr>
      <td>...</td>
      </tr>
      </table>")
    # Extract the "clean" table into a data frame 
    mountains <- mountains_html %>% 
      html_element("table#clean") %>% 
      html_table()

    mountains
    ```

5.  Read in `languages_raw_html`. Select all `div` *and* `p` elements in this HTML

<!-- -->

    ```{r}
    languages_raw_html <- read_html('<html> 
      <body> 
        <div>Python is perfect for programming.</div>
        <p>Still, R might be better suited for data analysis.</p>
        <small>(And has prettier charts, too.)</small>
      </body> 
    </html>')
    # Read in the HTML
    languages_html <- languages_raw_html
    # Select the div and p tags and print their text
    languages_html %>%
    	html_elements('div p') %>%
    	html_text()
    ```

6.  An html file has been read in for you with `read_html()` and is available through `structured_html.`Using `html_elements()`, find the shortest possible selector to select the first `div` in `structured_html`.

```{r}
    structured_html <- read_html("<html>
  <body>
    <div id = 'first'>
      <h1 class = 'big'>Joe Biden</h1>
      <p class = 'first blue'>Democrat</p>
      <p class = 'second blue'>Male</p>
    </div>
    <div id = 'second'>...</div>
    <div id = 'third'>
      <h1 class = 'big'>Donald Trump</h1>
      <p class = 'first red'>Republican</p>
      <p class = 'second red'>Male</p>
    </div>
  </body>
</html>")
    # Select the first div
    structured_html %>%
      html_elements('#first')
```

7.  An html file has been read in for you with `read_html()` and is available through `nested_html.`Select the last `p` node within the `div`.

```{r}
        nested_html<- read_html("<html>
          <body>
            <div>
              <p class = 'text'>A sophisticated text [...]</p>
              <p class = 'text'>Another paragraph following [...]</p>
              <p class = 'text'>Author: T.G.</p>
            </div>
            <p>Copyright: DC</p>
          </body>
        </html>")
        # This time for real: Select only the last node of the p's wrapped by the div
        nested_html  %>% 
        	html_elements('p.text:last-child')
```

8.  Here, your goal is to scrape a list (contained in the languages_html document) of all mentioned computer languages, but without the accompanying information in the sub-bullets.

```{r}
languages_html <- read_html("<ul id = 'languages'>
    <li>SQL</li>
    <ul>    
      <li>Databases</li>
      <li>Query Language</li>
    </ul>
    <li>R</li>
    <ul>
      <li>Collection</li>
      <li>Analysis</li>
      <li>Visualization</li>
    </ul>
    <li>Python</li>
  </ul>'")
```

For that you can use the `child`combinator "\>"

```{r}
# Extract only the text of the computer languages (without the sub lists)
languages_html %>% 
	html_elements('ul#languages > li') %>% 
	html_text2()
```
